# Distance, Knn, Cross Validation, and Generative Models

## Distance
*dist(A, B) = | A - B |*

```{r}
library(tidyverse)
library(dslabs)
library(caret)
set.seed(0)
if(!exists("mnist")) mnist <- read_mnist()
ind <- which(mnist$train$labels %in% c(2, 7)) %>% sample(500)
x <- mnist$train$images[ind,] # The predictor
y <- mnist$train$labels[ind]
y[1:3]
x_1 <- x[1,]
x_2 <- x[2,]
x_3 <- x[3,]
```


Distance between two numbers
```{r}
sqrt(sum((x_1 - x_2)^2))
sqrt(sum((x_1 - x_3)^2))
sqrt(sum((x_2 - x_3)^2))
```

Compute distance using matrix algebra
```{r}
sqrt(crossprod(x_1 - x_2))
sqrt(crossprod(x_1 - x_3))
sqrt(crossprod(x_2 - x_3))
```

Compute distance between each row
```{r}
d <- dist(x)
class(d)
as.matrix(d)[1:3,1:3]
```

visualize these distances
```{r}
image(as.matrix(d))
```

Order the distance by labels
```{r}
image(as.matrix(d)[order(y), order(y)])
```
Compute distance between predictors
```{r}
d <- dist(t(x))
dim(as.matrix(d))
d_492 <- as.matrix(d)[492,]
image(1:28, 1:28, matrix(d_492, 28, 28))
```

## K-nearest neighbors
- Estimates the conditional probabilities in a similar way to bin smoothing

- Easier to adapt to multiple dimensions

- For any point, we want an estimate of p(x1,x2), **look for the k nearest points to (x1,x2) and take an average of the 0s and 1s associated with these points**

**Neighborhood**: set of points used to compute the average.

A comparison
```{r}
library(caret)
```


- Logistic regression
```{r}
fit_glm <- glm(y~x_1+x_2, data=mnist_27$train, family="binomial")
p_hat_logistic <- predict(fit_glm, mnist_27$test)
y_hat_logistic <- factor(ifelse(p_hat_logistic > 0.5, 7, 2))
confusionMatrix(data = y_hat_logistic, reference = mnist_27$test$y)$overall[1]
```

- KNN model
```{r}
knn_fit <- knn3(y ~ ., data = mnist_27$train)
x <- as.matrix(mnist_27$train[,2:3])
y <- mnist_27$train$y
knn_fit <- knn3(x, y) # Look, as a matrix!
knn_fit <- knn3(y ~ ., data = mnist_27$train, k=5) # Y depends on all the predictors
y_hat_knn <- predict(knn_fit, mnist_27$test, type = "class")
confusionMatrix(data = y_hat_knn, reference = mnist_27$test$y)$overall["Accuracy"]
```

It's a very impressive improvement

Remember, logistic regression returns a line for the conditional probability. Now, KNN gives a more accurate result vs the ORIGINAL cond. prob.

```{r}
y_hat_knn <- predict(knn_fit, mnist_27$train, type = "class") 
confusionMatrix(data = y_hat_knn, reference = mnist_27$train$y)$overall["Accuracy"]
y_hat_knn <- predict(knn_fit, mnist_27$test, type = "class")  
confusionMatrix(data = y_hat_knn, reference = mnist_27$test$y)$overall["Accuracy"]
```


**Over-training**: The reason that we have higher accuracy in the train set compared to the test set.

- Its worst when we set *k = 1* (You're your own neighbor)
```{r}
knn_fit_1 <- knn3(y ~ ., data = mnist_27$train, k = 1)
y_hat_knn_1 <- predict(knn_fit_1, mnist_27$train, type = "class")
confusionMatrix(data=y_hat_knn_1, reference=mnist_27$train$y)$overall[["Accuracy"]]

```


**Over-smoothing**: *k* might be so large that it does not permit enough flexibility
```{r}
knn_fit_401 <- knn3(y ~ ., data = mnist_27$train, k = 401)
y_hat_knn_401 <- predict(knn_fit_401, mnist_27$test, type = "class")
confusionMatrix(data=y_hat_knn_401, reference=mnist_27$test$y)$overall["Accuracy"]
```


If we use the test set to pick this k, we should not expect the accompanying accuracy estimate to extrapolate to the real world **because we selected the *k* using the test set** 
```{r}
ks <- seq(3, 251, 2)
library(purrr)
accuracy <- map_df(ks, function(k){
    fit <- knn3(y ~ ., data = mnist_27$train, k = k)
    y_hat <- predict(fit, mnist_27$train, type = "class")
    cm_train <- confusionMatrix(data = y_hat, reference = mnist_27$train$y)
    train_error <- cm_train$overall["Accuracy"]
    y_hat <- predict(fit, mnist_27$test, type = "class")
    cm_test <- confusionMatrix(data = y_hat, reference = mnist_27$test$y)
    test_error <- cm_test$overall["Accuracy"]
    
tibble(train = train_error, test = test_error)
})

ks[which.max(accuracy$test)]
max(accuracy$test)
```

## k-fold Cross validation
- For **k -fold cross validation**, we divide the dataset into a training set and a test set. We train our algorithm exclusively on the training set and use the test set only for evaluation purposes. 

- For each set of algorithm parameters being considered, we want **an estimate of the MSE and then we will choose the parameters with the smallest MSE**. In *k*-fold cross validation, we randomly split the observations into  k  non-overlapping sets, and repeat the calculation for MSE for each of these sets. Then, we compute the average MSE and obtain an estimate of our loss. Finally, we can select the optimal parameter that minimized the MSE.

- In terms of how to select  k  for cross validation, **larger values of  k  are preferable** but they will also take much more computational time. For this reason, the choices of  **k=5  and  k=10**  are common.

## Bootstrap
```{r}
n <- 10^6
income <- 10^(rnorm(n, log10(45000), log10(3)))
qplot(log10(income), bins = 30, color = I("black"))

m <- median(income)
m
```

Now, the estimated data
```{r}
set.seed(1995, sample.kind="Rounding")
N <- 250
X <- sample(income, N)
M<- median(X)
M

library(gridExtra)
B <- 10^4
M <- replicate(B, {
    X <- sample(income, N)
    median(X)
})
p1 <- qplot(M, bins = 30, color = I("black"))
p2 <- qplot(sample = scale(M)) + geom_abline()
grid.arrange(p1, p2, ncol = 2)

mean(M)
sd(M)
```



- When we don't have access to the entire population, we can use **bootstrap** to estimate the population median *m*

- The bootstrap permits us to approximate a Monte Carlo simulation without access to the entire distribution. The general idea is relatively simple. We act as if the observed sample is the population. We then sample datasets (with replacement) of the same sample size as the original dataset. Then we compute the summary statistic, in this case the median, on this bootstrap sample.

```{r}
B <- 10^4
M_star <- replicate(B, {
    X_star <- sample(X, N, replace = TRUE)
    median(X_star)
})

tibble(monte_carlo = sort(M), bootstrap = sort(M_star)) %>%
    qplot(monte_carlo, bootstrap, data = .) + 
    geom_abline()

quantile(M, c(0.05, 0.95))
quantile(M_star, c(0.05, 0.95))

median(X) + 1.96 * sd(X) / sqrt(N) * c(-1, 1)

mean(M) + 1.96 * sd(M) * c(-1,1)

mean(M_star) + 1.96 * sd(M_star) * c(-1, 1)
```

## Generative Models

**Baye's rules**

- *Discriminative aoproaches*: Estimate the conditional probability directly and do not consider the distribution of the predictors.

- **Generative models** are methods that model the joint distribution and  X  (we model how the entire data, X and Y, are generated)

- p(x) = Pr(Y = 1 | X = x)

### Naive Bayes

```{r}
library(caret)
data("heights")
y <- heights$height
set.seed(2)
test_index <- createDataPartition(y, times = 1, p = 0.5, list = FALSE)
train_set <- heights %>% slice(-test_index)
test_set <- heights %>% slice(test_index)
```

Approximate the conditional distributions
```{r}
params <- train_set %>%
    group_by(sex) %>%
    summarize(avg = mean(height), sd = sd(height))
params
```

The prevalence
```{r}
pi <- train_set %>%
    summarize(pi = mean(sex == 'Female')) %>%
    .$pi
pi
```

Getting an actual rule
```{r}
x <- test_set$height
f0 <- dnorm(x, params$avg[2], params$sd[2])
f1 <- dnorm(x, params$avg[1], params$sd[1])
p_hat_bayes <- f1*pi / (f1*pi + f0*(1 - pi))
```

### Control prevalence (PI)

Computing sensitivity
```{r}
y_hat_bayes <- ifelse(p_hat_bayes > 0.5, "Female", "Male")
sensitivity(data = factor(y_hat_bayes), reference = factor(test_set$sex))
```

Computing specificity
```{r}
specificity(data = factor(y_hat_bayes), reference = factor(test_set$sex))
```

The Naive Bayes approach gives us a direct way to correct the imbalance between sensitivity and specificity by simply forcing  π^  to be whatever value we want it to be in order to better **balance specificity and sensitivity**.

Changing the cutoff of the decision rule
```{r}
p_hat_bayes_unbiased <- f1 * 0.5 / (f1 * 0.5 + f0 * (1 - 0.5))
y_hat_bayes_unbiased <- ifelse(p_hat_bayes_unbiased > 0.5, "Female", "Male")
sensitivity(data = factor(y_hat_bayes_unbiased), reference = factor(test_set$sex))
specificity(data = factor(y_hat_bayes_unbiased), reference = factor(test_set$sex))
```

Draw plot
```{r}
qplot(x, p_hat_bayes_unbiased, geom = "line") +
 geom_hline(yintercept = 0.5, lty = 2) +
 geom_vline(xintercept = 67, lty = 2)
```

### QDA & LDA
- **QDA**: Quadratic Discriminate Analysis

    - a version of Naive Bayes in which we assume that the distributions  pX|Y=1(x)  and  pX|Y=0(x)  are multivariate normal
    
    - an work well with a few predictors, but it becomes harder to use as the number of predictors increases
    
        - Once the number of parameters approaches the size of our data, the method becomes impractical due to overfitting
        
```{r}
# Load data
data("mnist_27")

# Estimate parameters from the data
params <- mnist_27$train %>%
 group_by(y) %>%
 summarize(avg_1 = mean(x_1), avg_2 = mean(x_2),
        sd_1 = sd(x_1), sd_2 = sd(x_2),
        r = cor(x_1, x_2))

# Contour plots
mnist_27$train %>% mutate(y = factor(y)) %>%
 ggplot(aes(x_1, x_2, fill = y, color = y)) +
 geom_point(show.legend = FALSE) +
 stat_ellipse(type="norm", lwd = 1.5)

# Fit model
library(caret)
train_qda <- train(y ~., method = "qda", data = mnist_27$train)
# Obtain predictors and accuracy
y_hat <- predict(train_qda, mnist_27$test)
confusionMatrix(data = y_hat, reference = mnist_27$test$y)$overall["Accuracy"]

# Draw separate plots for 2s and 7s
mnist_27$train %>% mutate(y = factor(y)) %>%
 ggplot(aes(x_1, x_2, fill = y, color = y)) +
 geom_point(show.legend = FALSE) +
 stat_ellipse(type="norm") +
 facet_wrap(~y)
```


- **LDA**: Linear Discriminate Analysis

    - Forcing the assumption that all predictors share the same standard deviations and correlations, the boundary will be a line, just as with logistic regression
    
    - the lack of flexibility does not permit us to capture the non-linearity in the true conditional probability function

```{r}
params <- mnist_27$train %>%
 group_by(y) %>%
 summarize(avg_1 = mean(x_1), avg_2 = mean(x_2),
        sd_1 = sd(x_1), sd_2 = sd(x_2),
        r = cor(x_1, x_2))
params <- params %>% mutate(sd_1 = mean(sd_1), sd_2 = mean(sd_2), r = mean(r))
train_lda <- train(y ~., method = "lda", data = mnist_27$train)
y_hat <- predict(train_lda, mnist_27$test)
confusionMatrix(data = y_hat, reference = mnist_27$test$y)$overall["Accuracy"]
```

# MOre than 3 classes
- Remember, **classes** means the output from *Y*

We obtain 3 probabilities for each data, and choose the biggest one. That will be our value.

Generative models can be very powerful, but only when we are able to **successfully approximate the joint distribution** of predictors conditioned on each class.

```{r}
if(!exists("mnist"))mnist <- read_mnist()

set.seed(3456)    #use set.seed(3456, sample.kind="Rounding") in R 3.6 or later
index_127 <- sample(which(mnist$train$labels %in% c(1,2,7)), 2000)
y <- mnist$train$labels[index_127] 
x <- mnist$train$images[index_127,]
index_train <- createDataPartition(y, p=0.8, list = FALSE)

# get the quadrants
# temporary object to help figure out the quadrants
row_column <- expand.grid(row=1:28, col=1:28)
upper_left_ind <- which(row_column$col <= 14 & row_column$row <= 14)
lower_right_ind <- which(row_column$col > 14 & row_column$row > 14)

# binarize the values. Above 200 is ink, below is no ink
x <- x > 200 

# cbind proportion of pixels in upper right quadrant and proportion of pixels in lower right quadrant
x <- cbind(rowSums(x[ ,upper_left_ind])/rowSums(x),
           rowSums(x[ ,lower_right_ind])/rowSums(x)) 

train_set <- data.frame(y = factor(y[index_train]),
                        x_1 = x[index_train,1],
                        x_2 = x[index_train,2])

test_set <- data.frame(y = factor(y[-index_train]),
                       x_1 = x[-index_train,1],
                       x_2 = x[-index_train,2])

train_set %>%  ggplot(aes(x_1, x_2, color=y)) + geom_point()
```

Let's train *QDA*, *LDA* & *KNN*
```{r}
train_qda <- train(y ~ ., method = "qda", data = train_set)
predict(train_qda, test_set, type = "prob") %>% head()
predict(train_qda, test_set) %>% head()
confusionMatrix(predict(train_qda, test_set), test_set$y)$table
confusionMatrix(predict(train_qda, test_set), test_set$y)$overall["Accuracy"]
train_lda <- train(y ~ ., method = "lda", data = train_set)
confusionMatrix(predict(train_lda, test_set), test_set$y)$overall["Accuracy"]
train_knn <- train(y ~ ., method = "knn", tuneGrid = data.frame(k = seq(15, 51, 2)),
    data = train_set)
confusionMatrix(predict(train_knn, test_set), test_set$y)$overall["Accuracy"]
train_set %>% mutate(y = factor(y)) %>% ggplot(aes(x_1, x_2, fill = y, color=y)) + geom_point(show.legend = FALSE) + stat_ellipse(type="norm")
```

