# Model Fitting and Recommendation Systems

Initial data
```{r}
library(dslabs)
mnist <- read_mnist()
class(mnist$train$labels)
table(mnist$train$labels)
```

Sample 10k rows
```{r}
set.seed(123)
index <- sample(nrow(mnist$train$images), 10000)
x <- mnist$train$images[index,]
y <- factor(mnist$train$labels[index])

index <- sample(nrow(mnist$test$images), 1000)
#note that the line above is the corrected code - code in video at 0:52 is incorrect
x_test <- mnist$test$images[index,]
y_test <- factor(mnist$test$labels[index])
```

## Pre-processing
We don't need a 0 variation feature

```{r}
library(matrixStats)
library(tidyverse)
library(caret)
library(dslabs)
sds <- colSds(x)
qplot(sds, bins = 256)

library(caret)
nzv <- nearZeroVar(x)
image(matrix(1:784 %in% nzv, 28, 28))

col_index <- setdiff(1:ncol(x), nzv)
length(col_index)
```

Remember, we are looking at pixels!

## Model Fitting for MNIST Data

Add name to columns
```{r}
colnames(x) <- 1:ncol(mnist$train$images)
colnames(x_test) <- colnames(x)

control <- trainControl(method = "cv", number = 10, p = .9)
train_knn <- train(x[,col_index], y,
                                method = "knn", 
                                tuneGrid = data.frame(k = c(1,3,5,7)),
                                trControl = control)
ggplot(train_knn)
```


Use KNN & select the best K
```{r}
n <- 1000
b <- 2
index <- sample(nrow(x), n)
control <- trainControl(method = "cv", number = b, p = .9)
train_knn <- train(x[index ,col_index], y[index],
                   method = "knn",
                   tuneGrid = data.frame(k = c(3,5,7)),
                   trControl = control)
fit_knn <- knn3(x[ ,col_index], y,  k = 3)

y_hat_knn <- predict(fit_knn,
                     x_test[, col_index],
                     type="class")
cm <- confusionMatrix(y_hat_knn, factor(y_test))
cm$overall["Accuracy"]

cm$byClass[,1:2]
```

Now, let's create forests of trees
```{r}
library(Rborist)
control <- trainControl(method="cv", number = 5, p = 0.8)
grid <- expand.grid(minNode = c(1,5) , predFixed = c(10, 15, 25, 35, 50))
train_rf <-  train(x[, col_index], y,
                   method = "Rborist",
                   nTree = 50,
                   trControl = control,
                   tuneGrid = grid,
                   nSamp = 5000)
ggplot(train_rf)
train_rf$bestTune
```

Tunning parameters
```{r}
fit_rf <- Rborist(x[, col_index], y,
                  nTree = 1000,
                  minNode = train_rf$bestTune$minNode,
                  predFixed = train_rf$bestTune$predFixed)

y_hat_rf <- factor(levels(y)[predict(fit_rf, x_test[ ,col_index])$yPred])
cm <- confusionMatrix(y_hat_rf, y_test)
cm$overall["Accuracy"]

rafalib::mypar(3,4)
for(i in 1:12){
     image(matrix(x_test[i,], 28, 28)[, 28:1], 
           main = paste("Our prediction:", y_hat_rf[i]),
           xaxt="n", yaxt="n")
}
```

## Variable importance

Lookinf at importance
```{r}
library(randomForest)
x <- mnist$train$images[index,]
y <- factor(mnist$train$labels[index])
rf <- randomForest(x, y,  ntree = 50)
imp <- importance(rf)
imp

image(matrix(imp, 28, 28))
```
There are many pixels that are not used, the yellow parts are the unused. We need to get rid of them

Now, let's see for our errors. 

```{r}
p_max <- predict(fit_knn, x_test[,col_index])
p_max <- apply(p_max, 1, max)
ind  <- which(y_hat_knn != y_test)
ind <- ind[order(p_max[ind], decreasing = TRUE)]
rafalib::mypar(3,4)
for(i in ind[1:12]){
    image(matrix(x_test[i,], 28, 28)[, 28:1],
                 main = paste0("Pr(",y_hat_knn[i],")=",round(p_max[i], 2),
                                                                        " but is a ",y_test[i]),
                 xaxt="n", yaxt="n")
}
```

The same for the random forests
```{r}
p_max <- predict(fit_rf, x_test[,col_index])$census  
p_max <- p_max / rowSums(p_max)
p_max <- apply(p_max, 1, max)
ind  <- which(y_hat_rf != y_test)
ind <- ind[order(p_max[ind], decreasing = TRUE)]
rafalib::mypar(3,4)
for(i in ind[1:12]){
    image(matrix(x_test[i,], 28, 28)[, 28:1], 
                 main = paste0("Pr(",y_hat_rf[i],")=",round(p_max[i], 2),
                               " but is a ",y_test[i]),
                 xaxt="n", yaxt="n")
}
```

## Ensambles
Using different methods to improve predictions

```{r}
p_rf <- predict(fit_rf, x_test[,col_index])$census
p_rf <- p_rf / rowSums(p_rf)
p_knn <- predict(fit_knn, x_test[,col_index])
p <- (p_rf + p_knn)/2
y_pred <- factor(apply(p, 1, which.max)-1)
confusionMatrix(y_pred, y_test)
```
## Recommendation Systems

Initial data
```{r}
library(dslabs)
library(tidyverse)
data("movielens")

head(movielens)

movielens %>%
     summarize(n_users = n_distinct(userId),
               n_movies = n_distinct(movieId))

keep <- movielens %>%
     dplyr::count(movieId) %>%
     top_n(5) %>%
     pull(movieId)
tab <- movielens %>%
     filter(userId %in% c(13:20)) %>% 
     filter(movieId %in% keep) %>% 
     select(userId, title, rating) %>% 
     spread(title, rating)
tab %>% knitr::kable()
```

User vs Movies
```{r}
users <- sample(unique(movielens$userId), 100)
rafalib::mypar()
movielens %>% filter(userId %in% users) %>% 
     select(userId, movieId, rating) %>%
     mutate(rating = 1) %>%
     spread(movieId, rating) %>% select(sample(ncol(.), 100)) %>% 
     as.matrix() %>% t(.) %>%
     image(1:100, 1:100,. , xlab="Movies", ylab="Users")
abline(h=0:100+0.5, v=0:100+0.5, col = "grey")
```

Movies rated
```{r}
movielens %>% 
     dplyr::count(movieId) %>% 
     ggplot(aes(n)) + 
     geom_histogram(bins = 30, color = "black") + 
     scale_x_log10() + 
     ggtitle("Movies")
```

Users that rate
```{r}
movielens %>%
     dplyr::count(userId) %>% 
     ggplot(aes(n)) + 
     geom_histogram(bins = 30, color = "black") + 
     scale_x_log10() +
     ggtitle("Users")
```

Create partition & RMSE function
```{r}
library(caret)
set.seed(755)
test_index <- createDataPartition(y = movielens$rating, times = 1,
                                  p = 0.2, list = FALSE)
train_set <- movielens[-test_index,]
test_set <- movielens[test_index,]

test_set <- test_set %>% 
     semi_join(train_set, by = "movieId") %>%
     semi_join(train_set, by = "userId")

RMSE <- function(true_ratings, predicted_ratings){
     sqrt(mean((true_ratings - predicted_ratings)^2))
}
```

## Building the Recommendation System

Y(u,i) = mu + bi + bu + epsilon(u, i)

```{r}
mu_hat <- mean(train_set$rating)
mu_hat

naive_rmse <- RMSE(test_set$rating, mu_hat)
naive_rmse

predictions <- rep(2.5, nrow(test_set))
RMSE(test_set$rating, predictions)

rmse_results <- data_frame(method = "Just the average", RMSE = naive_rmse)
```

Now, create the model
```{r}
fit <- lm(rating ~ as.factor(userId), data = movielens)
mu <- mean(train_set$rating) 
movie_avgs <- train_set %>% 
     group_by(movieId) %>% 
     summarize(b_i = mean(rating - mu))

movie_avgs %>% qplot(b_i, geom ="histogram", bins = 10, data = ., color = I("black"))

predicted_ratings <- mu + test_set %>% 
     left_join(movie_avgs, by='movieId') %>%
     .$b_i

model_1_rmse <- RMSE(predicted_ratings, test_set$rating)
rmse_results <- bind_rows(rmse_results,
                          data_frame(method="Movie Effect Model",
                                     RMSE = model_1_rmse ))

rmse_results %>% knitr::kable()

train_set %>% 
     group_by(userId) %>% 
     summarize(b_u = mean(rating)) %>% 
     filter(n()>=100) %>%
     ggplot(aes(b_u)) + 
     geom_histogram(bins = 30, color = "black")
```

We can use this: lm(rating ~ as.factor(movieId) + as.factor(userId))
For have a dependencie for Movies & Users, but this may crash our computer

Instead, we are going to:
```{r}
user_avgs <- test_set %>% 
     left_join(movie_avgs, by='movieId') %>%
     group_by(userId) %>%
     summarize(b_u = mean(rating - mu - b_i))

predicted_ratings <- test_set %>% 
     left_join(movie_avgs, by='movieId') %>%
     left_join(user_avgs, by='userId') %>%
     mutate(pred = mu + b_i + b_u) %>%
     .$pred

model_2_rmse <- RMSE(predicted_ratings, test_set$rating)
rmse_results <- bind_rows(rmse_results,
                          data_frame(method="Movie + User Effects Model",  
                                     RMSE = model_2_rmse ))
rmse_results %>% knitr::kable()
```

## Regularization

Penalty to low critics & very bad or good critics
```{r}
library(dslabs)
library(tidyverse)
library(caret)
data("movielens")
set.seed(755)
test_index <- createDataPartition(y = movielens$rating, times = 1,
                                  p = 0.2, list = FALSE)
train_set <- movielens[-test_index,]
test_set <- movielens[test_index,]
test_set <- test_set %>% 
     semi_join(train_set, by = "movieId") %>%
     semi_join(train_set, by = "userId")
RMSE <- function(true_ratings, predicted_ratings){
     sqrt(mean((true_ratings - predicted_ratings)^2))
}
mu_hat <- mean(train_set$rating)
naive_rmse <- RMSE(test_set$rating, mu_hat)
rmse_results <- data_frame(method = "Just the average", RMSE = naive_rmse)
mu <- mean(train_set$rating) 
movie_avgs <- train_set %>% 
     group_by(movieId) %>% 
     summarize(b_i = mean(rating - mu))
predicted_ratings <- mu + test_set %>% 
     left_join(movie_avgs, by='movieId') %>%
     .$b_i
model_1_rmse <- RMSE(predicted_ratings, test_set$rating)
rmse_results <- bind_rows(rmse_results,
                          data_frame(method="Movie Effect Model",
                                     RMSE = model_1_rmse ))
user_avgs <- test_set %>% 
     left_join(movie_avgs, by='movieId') %>%
     group_by(userId) %>%
     summarize(b_u = mean(rating - mu - b_i))
predicted_ratings <- test_set %>% 
     left_join(movie_avgs, by='movieId') %>%
>     left_join(user_avgs, by='userId') %>%
     mutate(pred = mu + b_i + b_u) %>%
     .$pred
model_2_rmse <- RMSE(predicted_ratings, test_set$rating)
rmse_results <- bind_rows(rmse_results,
                          data_frame(method="Movie + User Effects Model",  
                                     RMSE = model_2_rmse ))

test_set %>% 
     left_join(movie_avgs, by='movieId') %>%
     mutate(residual = rating - (mu + b_i)) %>%
     arrange(desc(abs(residual))) %>% 
     select(title,  residual) %>% slice(1:10) %>% knitr::kable()

movie_titles <- movielens %>% 
     select(movieId, title) %>%
     distinct()
movie_avgs %>% left_join(movie_titles, by="movieId") %>%
     arrange(desc(b_i)) %>% 
     select(title, b_i) %>% 
     slice(1:10) %>%  
     knitr::kable()

movie_avgs %>% left_join(movie_titles, by="movieId") %>%
     arrange(b_i) %>% 
     select(title, b_i) %>% 
     slice(1:10) %>%  
     knitr::kable()

train_set %>% dplyr::count(movieId) %>% 
     left_join(movie_avgs) %>%
     left_join(movie_titles, by="movieId") %>%
     arrange(desc(b_i)) %>% 
     select(title, b_i, n) %>% 
     slice(1:10) %>% 
     knitr::kable()

train_set %>% dplyr::count(movieId) %>% 
     left_join(movie_avgs) %>%
     left_join(movie_titles, by="movieId") %>%
     arrange(b_i) %>% 
     select(title, b_i, n) %>% 
     slice(1:10) %>% 
     knitr::kable()
lambda <- 3
mu <- mean(train_set$rating)
movie_reg_avgs <- train_set %>% 
     group_by(movieId) %>% 
     summarize(b_i = sum(rating - mu)/(n()+lambda), n_i = n()) 

data_frame(original = movie_avgs$b_i, 
           regularlized = movie_reg_avgs$b_i, 
           n = movie_reg_avgs$n_i) %>%
     ggplot(aes(original, regularlized, size=sqrt(n))) + 
     geom_point(shape=1, alpha=0.5)

train_set %>%
     dplyr::count(movieId) %>% 
     left_join(movie_reg_avgs) %>%
     left_join(movie_titles, by="movieId") %>%
     arrange(desc(b_i)) %>% 
     select(title, b_i, n) %>% 
     slice(1:10) %>% 
     knitr::kable()

train_set %>%
     dplyr::count(movieId) %>% 
     left_join(movie_reg_avgs) %>%
     left_join(movie_titles, by="movieId") %>%
     arrange(b_i) %>% 
     select(title, b_i, n) %>% 
     slice(1:10) %>% 
     knitr::kable()

predicted_ratings <- test_set %>% 
     left_join(movie_reg_avgs, by='movieId') %>%
     mutate(pred = mu + b_i) %>%
     .$pred

model_3_rmse <- RMSE(predicted_ratings, test_set$rating)
rmse_results <- bind_rows(rmse_results,
                          data_frame(method="Regularized Movie Effect Model",  
                                     RMSE = model_3_rmse ))
rmse_results %>% knitr::kable()

lambdas <- seq(0, 10, 0.25)
mu <- mean(train_set$rating)
just_the_sum <- train_set %>% 
     group_by(movieId) %>% 
     summarize(s = sum(rating - mu), n_i = n())
rmses <- sapply(lambdas, function(l){
     predicted_ratings <- test_set %>% 
          left_join(just_the_sum, by='movieId') %>% 
          mutate(b_i = s/(n_i+l)) %>%
          mutate(pred = mu + b_i) %>%
          .$pred
     return(RMSE(predicted_ratings, test_set$rating))
})
qplot(lambdas, rmses)  
lambdas[which.min(rmses)]

lambdas <- seq(0, 10, 0.25)
rmses <- sapply(lambdas, function(l){
     mu <- mean(train_set$rating)
     b_i <- train_set %>%
          group_by(movieId) %>%
          summarize(b_i = sum(rating - mu)/(n()+l))
     b_u <- train_set %>% 
          left_join(b_i, by="movieId") %>%
          group_by(userId) %>%
          summarize(b_u = sum(rating - b_i - mu)/(n()+l))
     predicted_ratings <- 
          test_set %>% 
          left_join(b_i, by = "movieId") %>%
          left_join(b_u, by = "userId") %>%
          mutate(pred = mu + b_i + b_u) %>%
          .$pred
     return(RMSE(predicted_ratings, test_set$rating))
})

qplot(lambdas, rmses)  

lambda <- lambdas[which.min(rmses)]
lambda

rmse_results <- bind_rows(rmse_results,
                          data_frame(method="Regularized Movie + User Effect Model",  
                                     RMSE = min(rmses)))
rmse_results %>% knitr::kable()
```






